Dissertation Proposal

### 2023-11-07 proposal.tex
RQ1

\subsection{Multiple Populations}%
\label{s:populations}

@@

\subsection{Find Misconceptions in Student-written Programs}%
\label{s:beyond-experts-blindsopts}

To find a reasonably comprehensive collection of misconceptions, I
seek for inputs from students. Most prior works
(\Cref{s:background-misconceptions}) have experts coming up with a
list of potential misconceptions and then filter the potential
misconceptions with student inputs. This approach can't find
misconceptions not envisioned by experts and hence is hindered by
experts' blind spots. To overcome such bias, I analyze a University~1
dataset produced by a two-year process to find (a) student-written
programs that tend to trip up students and (b) incorrect responses to
those programs (\Cref{s:gen-prob-quizius}), using a tool for the
purpose (\Cref{s:quizius}). I have been cleaning up those programs
(\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
these programs are still tricky in the same way
(\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
\Cref{s:misconceptions-tutor} presents the final list of
misconceptions found with this process.

\subsection{Find Misconceptions with Expert-written Programs}

I have another source of misconceptions. I have been creating a tutor
to answer my RQ2. The tutor, which I explain in details when I talk
about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks students
to predict the result(s) of running a program. I have found several
unanticipated patterns of errors (\Cref{s:misconceptions-tutor}),
which suggest the existence of unknown misconceptions. The tutor is
not collecting students' rationale for their answers. So I do not know
what are the underlying misconceptions (although some of the wrong
answers are somewhat self-explanatory). I plan to change the tutor
such that it asks students' for their explanations when they give a
wrong answer. I plan to eyeball the explanations to find new
misconceptions and to further confirm that anticipated wrong answers
do represent the expected misconceptions.

\subsection{Multiple-Syntax}

[TODO]

>>>>

\subsection{Populations}%
\label{s:populations}

@@

I am contacting professors from various other universities. Several of
them have expressed interests in using my instruments. So the final
results will likely include more populations.

\subsection{Find Misconceptions in Student-written Programs}%
\label{s:find-misconceptions-with-quizius}

To find a reasonably comprehensive collection of misconceptions, I
seek for inputs from students. Most prior works
(\Cref{s:background-misconceptions}) have experts coming up with a
list of potential misconceptions and then filter the potential
misconceptions with student inputs. This approach can't find
misconceptions not envisioned by experts and hence is hindered by
experts' blind spots.

To overcome such bias, I analyze a University~1 dataset produced by a
two-year process to find (a) student-written programs that tend to
trip up students and (b) incorrect responses to those programs
(\Cref{s:gen-prob-quizius}), using a tool for the purpose
(\Cref{s:quizius}). I have been cleaning up those programs
(\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
these programs are still tricky in the same way
(\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
\Cref{s:misconceptions-tutor} presents the final list of
misconceptions found with this process.

\subsection{Find Misconceptions with Expert-written Programs}%
\label{s:find-misconceptions-with-tutor}

I have another source of misconceptions. I have been creating a tutor
to answer my RQ2, which I explain in details in \Cref{s:plan-rq2}. The
tutor includes many tasks that asks students to predict the result(s)
of running a program. Although these programs originate from
student-written programs (\Cref{s:find-misconceptions-with-quizius}), they
have been modified to fit tutoring purpose. Perhaps because of the
modification or because of the population variance, I have found
several unanticipated patterns of errors
(\Cref{s:misconceptions-tutor}), which suggest unknown misconceptions.

I plan to change the tutor such that it asks students' to explain
their prediction. I plan to eyeball the explanations to find new
misconceptions. This piece of data also helps me to confirm that
anticipated wrong answers do represent the expected misconceptions.

\subsection{Reduce Syntactic Bias}

The processes described in
\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
potentially introduce biases to the found misconceptions. For example,
current results find that some students think the following program
produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
\begin{lstlisting}
  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}
This checks whether $\verb|3| > \verb|n|$, but those students
presumably vocalized it as ``greater than \lstinline|3|,
\lstinline|n|?''. Although I have tried to remove programs like this
from my instruments, there is still a risk of having biase in student
errors. I plan to adjust the process descrbed in
\Cref{s:find-misconceptions-with-tutor} such that it presents programs
in multiple languages: the SMoL version, a JavaScript translation, and
a Python version.
